---
title: "Time Series Project"
output: pdf_document
---

# Import libraries

```{r }
library(fpp3)
library(fabletools)
library(zoo)
library(tseries)
library(quantmod) # download data form Yahoo finance
library(moments) # to know summary statistics of data
library(tinytex)
library(forecast)
library(tsintermittent)
library(expsmooth)
library(readxl)
library(dplyr)
library(tsibble)
library(ggplot2)
library(fable.prophet) 
library(prophet)
library(fable.prophet) 
#install.packages("fable.prophet")
#install.packages("expsmooth")
```

# Read Data

```{r }

##*************************************************##
fred_data <- read_excel("inflation_project.xls") 

# Create date variable
inflation <- zoo::fortify.zoo(fred_data)

inflation <- as_tsibble(inflation, index = Index) # create tissble object

inflation <- inflation |>
mutate(month = row_number()) |>
update_tsibble(index = month, regular = TRUE)
head(inflation)
tail(inflation)

# Plot of CPI only
inflation %>% autoplot(cpi,colour=("blue")) +
labs(y="% Change From Year Ago",
     x="Months after 1/1/1972",
     title = "Consumer Price Index for Canada")
  
# All graphs together
inflation %>% 
  pivot_longer(c(cpi, exchange_rate, unemp_rate), names_to="Series") %>%
autoplot(value) +
labs(y="% Change From Year Ago",
     x="Months after 1/1/1972",
     title = "CPI, Exchange Rate and Unemployment Rate: Canada")

# Summary of the data
summary(inflation)

#Missing data check
tsoutliers(inflation$cpi)
# No outliers

tsoutliers(inflation$exchange_rate)
# Replace exchange_rate outliers
inflation$exchange_rate[442]<-round(3.922876)
inflation$exchange_rate[443]<-round(4.781721)
inflation$exchange_rate[444]<-round(5.640567)
inflation$exchange_rate[445]<-round(6.499412)
inflation$exchange_rate[446]<-round(7.358258)
inflation$exchange_rate[447]<-round(8.217103)
inflation$exchange_rate[448]<-round(9.075949)
inflation$exchange_rate[449]<-round(9.934794)

tsoutliers(inflation$unemp_rate)
# Replace unemp_rate outliers
inflation$unemp_rate[126]<-round(43.632917)
inflation$unemp_rate[127]<-round(44.210275)
inflation$unemp_rate[128]<-round(44.787633)
inflation$unemp_rate[129]<-round(45.364990)
inflation$unemp_rate[130]<-round(45.942347)
inflation$unemp_rate[131]<-round(46.519705)
inflation$unemp_rate[132]<-round(47.097062)
inflation$unemp_rate[450]<-round(37.762417)
inflation$unemp_rate[451]<-round(36.815155)
inflation$unemp_rate[452]<-round(35.867893)
inflation$unemp_rate[579]<-round(-2.324115)
inflation$unemp_rate[580]<-round(-2.924089)
inflation$unemp_rate[581]<-round(-3.524064)
inflation$unemp_rate[582]<-round(-4.124038)
inflation$unemp_rate[583]<-round(-4.724013)
inflation$unemp_rate[584]<-round(-5.323988)
inflation$unemp_rate[585]<-round(-5.923962)
inflation$unemp_rate[586]<-round(-6.523937)
inflation$unemp_rate[587]<-round(-7.123912)
inflation$unemp_rate[588]<-round(-7.723886)
inflation$unemp_rate[589]<-round(-8.323861)
inflation$unemp_rate[590]<-round(-8.923835)
inflation$unemp_rate[592]<-round(-15.097402)
inflation$unemp_rate[593]<-round(-20.670995)
inflation$unemp_rate[594]<-round(-26.244588)
inflation$unemp_rate[133]<-round(44.84831)
##*************************************************##
# Train Dataset
train_inflation <- inflation %>% filter(Date<='2022-08-31')
head(train_inflation)
tail(train_inflation)

# Test Dataset
test_inflation <- inflation %>% filter(Date>'2022-08-31' & Date<='2023-08-31')


# Sign Correlation
rho.cal<-function(X){
  rho.hat<- cor(X-mean(X), sign(X-mean(X)))
  return(as.numeric(rho.hat))
}

# Sign Correlation
rho.cal(inflation$cpi)
##*************************************************##

```

* The distribution is normal.


# Benchmark methods

```{r}
# benchmark models (mean, naive, drift)
inflation_ben_fit <- train_inflation |>
  model(
    Mean = MEAN(cpi),
    Naïve = NAIVE(cpi),
    Drift = RW(cpi ~ drift())
  )

# forecast
inflation_ben_fc <- inflation_ben_fit |>
  forecast(h=12)

# Plot the forecasts with whole data
inflation_ben_fc |>
  autoplot(inflation)+ 
  labs(y = "% Change From Year Ago",
       x = "Months after 1/1/1972",
       title = "CPI Index Forecast : Canada",
       subtitle = "(Jan 1972 - Aug 2023)") 
  #+guides(colour = guide_legend(title = "Forecast"))
 
#Plot only forecast period        
inflation_ben_fc |>
  autoplot(test_inflation) + 
  labs(y = "% Change From Year Ago",
        x = "Months after 1/1/1972",
       title = "CPI Index Forecast : Canada",
       subtitle = "(Sept 2022 - Aug 2023)")

#Accuracy()
accuracy.ben <-  accuracy(inflation_ben_fc,inflation)

#Residual check
train_inflation |>model(NAIVE(cpi)) |>gg_tsresiduals()

#Portmanteau tests for autocorrelation
# Test for Naive method
t<- train_inflation |>
  model(NAIVE(cpi)) |>
  augment()
t |> features(.innov, box_pierce, lag = 25)

# Test for drift Method
fit <- train_inflation |> model(RW(cpi ~ drift()))
tidy(fit)
augment(fit) |> features(.innov, ljung_box, lag=25)

```

* Mean method gave very bad forecast.
* Residuals for Naive method are not white noise . However, residuals for Drift method are white noise.

# ARIMA Model

```{r}
fit__cpi_ar <- train_inflation |>
  model(
    arima = ARIMA(cpi)
    #ets = ETS(CPI_Index)
    #prophet = prophet(cpi)
  )
report(fit__cpi_ar) 

#Residual check
fit__cpi_ar |> gg_tsresiduals()
#Portmanteau tests for autocorrelation
augment(fit__cpi_ar) |> features(.innov, ljung_box, dof = 1, lag = 50)

# Plot the forecasts with whole data
fit__cpi_ar_fc <- fit__cpi_ar  %>% forecast(h = 12) 
fit__cpi_ar_fc |>  autoplot(inflation)+ 
  labs(y = "% Change From Year Ago",
       x = "Months after 1/1/1972",
       title = "CPI Index Forecast : Canada",
       subtitle = "(Jan 1972 - Aug 2023)") 

#Plot only forecast period        
fit__cpi_ar_fc %>% autoplot(test_inflation)+ 
  labs(y = "% Change From Year Ago",
        x = "Months after 1/1/1972",
       title = "CPI Index Forecast : Canada",
       subtitle = "(Sept 2022 - Aug 2023)")

#Accuracy
accuracy.ar <- accuracy(fit__cpi_ar_fc,inflation)

```

* Residuals are not white noise..

# Bootstrap

```{r}
fit <- train_inflation |>
  model(ARIMA(cpi))
sim <- fit |> generate(h = 12, times = 5, bootstrap = TRUE) # Each time 30 forecasts
head(sim)


train_inflation |>
  ggplot(aes(x = month)) +
  geom_line(aes(y = cpi)) +
  geom_line(aes(y = .sim, colour = as.factor(.rep)),
    data = sim) +
  labs(title="Five simulated sample paths based on naive method with boostrapped residuls", y="$US" ) +
  guides(colour = "none")


fc <- fit |> forecast(h = 12, bootstrap = TRUE) 
fc %>% 
autoplot(test_inflation) +
  labs(title="Forecasts of CIP")

#accuracy(GOOG_fc, GOOG_stock) %>% select(.model, RMSE:MAPE)
```


# Neural Network model

```{r }
# NNAR Model
fit_cpi_nn <- train_inflation %>% model(NNETAR(cpi))

#Residual check
fit_cpi_nn |> gg_tsresiduals()
#Portmanteau tests for autocorrelation
augment(fit_cpi_nn) |> features(.innov, ljung_box, lag = 50)

# Plot the forecasts with whole data
fit_cpi_fc <- fit_cpi_nn %>% forecast(h = 12) 
fit_cpi_fc %>%  autoplot(inflation)+ 
  labs(y = "% Change From Year Ago",
       x = "Months after 1/1/1972",
       title = "CPI Index Forecast : Canada",
       subtitle = "(Jan 1972 - Aug 2023)") 

#Plot only forecast period        
fit_cpi_fc %>% autoplot(test_inflation)+ 
  labs(y = "% Change From Year Ago",
        x = "Months after 1/1/1972",
       title = "CPI Index Forecast : Canada",
       subtitle = "(Sept 2022 - Aug 2023)")

# Accuracy
accuracy.nn <- accuracy(fit_cpi_fc,inflation)
```
* * Residuals are white noise.

# Timeseries Regression Model

```{r}
#Standard Time Series Regression Model
fit_stand <- train_inflation %>%
  model(lm = TSLM(cpi ~ exchange_rate+ unemp_rate))
report(fit_stand)

#***************** Scenario Based Forecasting**********************#
# It is assumed that exchange rate and unemployment rate will remain constant 
# for 12 months(testing period).
cpi_future <- new_data( train_inflation, 12) |>
  mutate(
 exchange_rate = 2.8 ,
    unemp_rate = -10
  )
#******************************************************************#

# Plot the forecasts with whole data
fit_stand_fc <-forecast(fit_stand, cpi_future) 
fit_stand_fc |>
  autoplot(inflation) + 
  labs(y = "% Change From Year Ago",
       x = "Months after 1/1/1972",
       title = "CPI Index Forecast : Canada",
       subtitle = "(Jan 1972 - Aug 2023)") 

#Plot only forecast period  
forecast(fit_stand, cpi_future) |>
  autoplot(test_inflation) + 
  labs(y = "% Change From Year Ago",
        x = "Months after 1/1/1972",
       title = "CPI Index Forecast : Canada",
       subtitle = "(Sept 2022 - Aug 2023)")

#Residual check
fit_stand |> gg_tsresiduals()
#Portmanteau tests for autocorrelation
augment(fit_stand) |> features(.innov, ljung_box, lag = 50)

# Accuracy
accuracy.sr <- accuracy(fit_stand_fc ,inflation)
```

# Dynamic Regression Model

```{r}
#Dynamic Regression Model with ARIMA errors
fit_dyn <- train_inflation |>
  model(ARIMA(cpi ~ exchange_rate+ unemp_rate))
report(fit_dyn )

# Plot the forecasts with whole data
fit_dyn_fc <- forecast(fit_dyn, cpi_future) 

fit_dyn_fc |>
  autoplot(inflation) + 
  labs(y = "% Change From Year Ago",
       x = "Months after 1/1/1972",
       title = "CPI Index Forecast : Canada",
       subtitle = "(Jan 1972 - Aug 2023)") 

#Plot only forecast period  
forecast(fit_dyn, cpi_future) |>
  autoplot(test_inflation) + 
  labs(y = "% Change From Year Ago",
        x = "Months after 1/1/1972",
       title = "CPI Index Forecast : Canada",
       subtitle = "(Sept 2022 - Aug 2023)")

#Residual check
fit_dyn |> gg_tsresiduals()
augment(fit_dyn) |> features(.innov, ljung_box, dof=1, lag = 50)

# Accuracy
accuracy.dr <- accuracy(fit_dyn_fc,inflation)
```

# Hybrid Model

```{r}
# Combine neural network model and dynamic regression model
combine1 <- train_inflation %>% model(
  Dynamic = ARIMA(cpi ~ exchange_rate+ unemp_rate),
  Neural_Network =  NNETAR((cpi))
) %>% 
mutate( Combination=(Dynamic+ Neural_Network)/2 ) 


fc_combine <- forecast(combine1, cpi_future) 

# Plot the forecasts with whole data
fc_combine %>%  autoplot(inflation)+ 
  labs(y = "% Change From Year Ago",
       x = "Months after 1/1/1972",
       title = "CPI Index Forecast : Canada",
       subtitle = "(Jan 1972 - Aug 2023)") 

#Plot only forecast period  
fc_combine %>%  autoplot(test_inflation)+ 
  labs(y = "% Change From Year Ago",
        x = "Months after 1/1/1972",
       title = "CPI Index Forecast : Canada",
       subtitle = "(Sept 2022 - Aug 2023)")


#Accuracy
accuracy.hm <- accuracy(fc_combine,inflation)
```

# Accuracy of all Models
```{r}
accuracy.table <- rbind(accuracy.ben, accuracy.ar,accuracy.nn,accuracy.sr,accuracy.dr,accuracy.hm)
accuracy.table
```


**************************************************************************************************************************

